Discussion

### 2020-08-15 text/discussion.Rmd

We attempted to account for our clusters in terms of more trivial explanations.
For example, it could be that the classes we identified are consistent with a
signature of overall disease progression, such that successive, adjacent classes
on the UMAP projection exhibit increasing specificity. We identified no such
trend in the number of mutations (see **figure 4**), the relative staging
(**figure 2D**) or metastasis (**figure 2E**); and each class instead was
associated with unique combinations of pathways (as in **figures 5B, C**).

@@

By modeling CDC longevity data as a baseline risk function we showed that each
of our pathway disruption classes exhibit cancer-type specific effects on
expected survival (**figure 7**). Contrary to what one might expect at the
outset, we found that models assuming different survival based on class and
cancer type fit better than models in which they are independent
factors. However, considering that within each cancer type there are different
clinical standards of care, and even within classes of drugs the preferred
treatment can very between cancers, it makes sense that we observe tissue-specific class effects. Contrast the situation with ovarian _vs._ breast cancer, which
are both hormonally driven cancers, for example. Ovarian cancer has but one main
treatment axis, platinum, whereas breast cancer patients have a variety of
treatment regimens based on molecular subtype and other factors. It is easy to
see how standard of care for one cancer and class could produce different
outcomes than the same class of a different cancer type. Unfortunately, given
the diversity drug classes and treatments (including non-chemical treatments
such as surgery and palliative care), we lack sufficient power 
to explore these variables in the TCGA data. It is our hope that 
future studies will help to distinguish between treatment-specific
effects on survival given different pathway disruption classes.

>>>>

We attempted to account for our clusters in terms of more trivial explanations.
For example, it could be that the classes we identified are consistent with a
signature of overall disease progression, such that successive, adjacent classes
on the UMAP projection exhibit increasing specificity. We identified no such
trend in the number of mutations (see **figure 4**), the relative staging
(**figure 3D**) or metastasis (**figure 3E**); and each class instead was
associated with unique combinations of pathways (as in **figures 5B, C**).

@@

By modeling CDC longevity data as a baseline risk function we showed that each
of our pathway disruption classes exhibit cancer-type specific effects on
expected survival (**figure 9**). Contrary to what one might expect at the
outset, we found that models assuming different survival based on class and
cancer type fit better than models in which they are independent
factors. However, considering that within each cancer type there are different
clinical standards of care, and even within classes of drugs the preferred
treatment can very between cancers, it makes sense that we observe tissue-specific class effects. Contrast the situation with ovarian _vs._ breast cancer, which
are both hormonally driven cancers, for example. Ovarian cancer has but one main
treatment axis, platinum, whereas breast cancer patients have a variety of
treatment regimens based on molecular subtype and other factors. It is easy to
see how standard of care for one cancer and class could produce different
outcomes than the same class of a different cancer type. Unfortunately, given
the diversity drug classes and treatments (including non-chemical treatments
such as surgery and palliative care), we lack sufficient power 
to explore these variables in the TCGA data. It is our hope that 
future studies will help to distinguish between treatment-specific
effects on survival given different pathway disruption classes.
